---
title: "PSU Womens Volleyball"
author: "Lily Mayer, Ayla Orona, Kate Miller"
date: "2023-11-29"
output: html_document
---

All code should be written according to a Style Guide of your choice. List this Style Guide as a code comment in your first code chunk.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# List Style Guide in this code chunk
library(rvest)
library(dplyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(stringr)
library(tidyr)
```

```{r espnTable}
espnPSUTables <- read_html(
  x = "link"
  # I haven't found any ESPN related items, so we may have to rethink where we get data from.
) %>%
  html_elements(css = "table") %>%
  html_table()
View(espnPSUTables)
```

```{r PSURosterTable}
psuVBRoster <- read_html(
  x = "https://gopsusports.com/sports/womens-volleyball/roster"
) %>%
  html_elements(css = "table") %>%
  html_table()
psuVBRoster <- bind_cols(psuVBRoster[[3]])
#saves correct data frame from harvest

psuVBRoster <- psuVBRoster %>%
  tidyr::separate_wider_delim(
    cols = `Ht.`,
    delim = '-',
    names = c('Feet', 'Inches')
  ) %>%
  #separates height column into feet and inches columns
  mutate_at(vars('Feet', 'Inches'), as.numeric) %>%
  #turns feet and inches into numeric types
  mutate('Height (in.)' = Inches + (Feet * 12)) %>%
  #creates a height column in inches
  select('#', 'Full Name', 'Pos.', 'Elig.', 'Height (in.)', 'Hometown / High School')
#reorders column and removes feet and inches columns

View(psuVBRoster)
```

```{r PSURosterStatsTable}
psuRosterStats <- read_html(
  x = 'https://gopsusports.com/sports/womens-volleyball/stats'
) %>%
  html_elements(css = 'table') %>%
  html_table()
#harvests the player data from the gopsusports website

psuRosterOffensiveHeading <- bind_cols(psuRosterStats[[2]][1, ])
#saves the column headers as a single row

psuRosterOffensiveStats <- bind_cols(psuRosterStats[[2]][2:nrow(psuRosterStats[[2]]) , ])
#saves the rest of the player data as a dataframe

colnames(psuRosterOffensiveStats) <- c(psuRosterOffensiveHeading[1, ])
#inserts column headers into player data

psuOffensive <- psuRosterOffensiveStats %>%
  select(!contains('Link'))
#removes the 'Bio Link' column

psuOffensive <- psuOffensive[1:17, ]
#removes the 'Total' and 'Opponent' rows

psuOffensive$Player <- str_sub(psuOffensive$Player, end = -25)
#removes the repeated names from the player column

psuOffensive <- psuOffensive %>%
  tidyr::separate_wider_delim(
    cols = Player,
    delim = ',',
    names = c('Last', 'First')
  ) %>%
  #Separates the player names into separate columns for first and last name
  mutate(
    'Player' = paste(First, Last, sep=' '), 
    .before = First
  ) %>%
  #concatenates the first and last names together
  mutate(
    `Player` = stringr::str_squish(`Player`)) %>%
  select(!'First') %>%
  select(!'Last')
  #removes excess whitespace and removes individual name columns

psuOffensive <- psuOffensive %>%
  mutate_at(vars(!Player), as.numeric)
#turns all columns except Player into numeric types

psuOffensive <- inner_join(
  x = psuOffensive, 
  y = psuVBRoster, 
  by = join_by('Player' == 'Full Name', '#' == '#')
)
#joins the roster data into the offensive statistics

View(psuOffensive)
```

## Research Questions
State your research questions and explain them.

## Read in data and perform necessary wrangling and cleaning
Describe your data sources (where they come from, what was their original purpose, who/what comprise the cases in each).

Describe what attributes you'll focus your analysis on (mention if they are part of your data sets or if you created them out of your data sets).

## Conduct Exploratory Data Analysis
All outputs should be accompanied by narrative text to explain what the reader should be seeing.

Create multiple data visualizations that assist both the team and readers in understanding the data.
Data visualizations should show a variety of your skills and geometries.

Create at least one visual table (not a display of raw data) that assists both the team and readers in understanding the data.

Have narrative text explaining every data visualization and table as well as setting a framework.
