---
title: "Exploring PSU Women's Volleyball and OSU Women's Volleyball Data"
author: "Lily Mayer, Ayla Orona, Kate Miller"
date: "2023-11-29"
output: html_document
---
# Introduction

The following document aims to provide details on Penn State Women's Volleyball data.

### The General Rules of Collegiate Volleyball

Volleyball is a game that consists of two teams each with six players hitting a ball over a net to score points. Each play starts by a setter on one team hitting the ball over the net. After the ball is served, each team is allowed to touch the ball three times before the ball must go over the net. Throughout the game, when the ball hits the ground of an opposing teams side, points are given to the team that last hit the ball. 

### Volleyball Terms Used in This Document

The following column headings were collected from the Penn State Women's Volleyball website and may be used in the data frames, tables, and visualizations. Each are based on individual volleyball players.  
  
Player - Penn State Women's Volleyball player first and last name  
SP - Sets played  
MP -  
MS -  
PTS - Total points earned for the 2023 season  
PTS/S -  
K - Total number of kills  
K/S -   
E -  
TA -  
PCT -   
A -   
A/S -   
SA -  
SA/S -   
SE -   
  
The following are the different positions on the volleyball team, which may be referenced in a data visualization or table.  
  
MB -   
L -   
DS -   
OH -   
S -   
RS -   
=======
The following column headings were collected from the Penn State Women's Volleyball website and may be used in the data frames, tables, and visualizations. Each are based on individual volleyball players.

Player - Penn State Women's Volleyball player first and last name
SP - Total number of sets played
MP - Total number of matches played
MS - Total number of matches where the player started
PTS - Total points earned for the 2023 season
PTS/S - Average number of points earned per sets played
K - Total number of kills ****
K/S - Average number of kills executed per sets played
E - Total number of errors
TA - Total attacks ***
PCT - Attack percentage ***
A - Total number of assists where the player threw the ball to a teammate who then won the point
A/S - Average number of assists per sets played
SA - Total number of service aces, an ace means the serving team won the point before the receiving team could return the ball over the net
SA/S - Average number of service aces per sets played
SE - Total number of service errors, meaning the serve did not throw the ball over the net or threw it out of bounds

The following are the different positions on the volleyball team, which may be referenced in a data visualization or table.

MB - Middleblocker
L - Libero
DS - Defensive Specialist
OH - Outside Hitter
S - Setter
RS - Rightside Hitter
>>>>>>> 4eeac16cd7f0ba04dbceb6a2147ad639e9fee559

Some players have more than one position. In these cases, the player's data will be included twice, once for each position, for comparisons that relate to the player's position. 

### Research Questions


All code should be written according to a Style Guide of your choice. List this Style Guide as a code comment in your first code chunk.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
# List Style Guide in this code chunk
#BOAST STYLE GUIDE OR GOOGLE'S R GUIDE
library(rvest)
library(ggplot2)
library(knitr)
library(kableExtra)
library(stringr)
library(tidyr)
library(pdftools)
library(tidyverse)
library(esquisse)
library(janitor)
```

```{r PSURosterTable}
psuVBRoster <- read_html(
  x = "https://gopsusports.com/sports/womens-volleyball/roster"
) %>%
  html_elements(css = "table") %>%
  html_table()
psuVBRoster <- bind_cols(psuVBRoster[[3]])
#saves correct data frame from harvest

psuVBRoster <- psuVBRoster %>%
  tidyr::separate_wider_delim(
    cols = `Ht.`,
    delim = '-',
    names = c('Feet', 'Inches')
  ) %>%
  #separates height column into feet and inches columns
  mutate_at(vars('Feet', 'Inches'), as.numeric) %>%
  #turns feet and inches into numeric types
  mutate('Height (in.)' = Inches + (Feet * 12)) %>%
  #creates a height column in inches
  select('#', 'Full Name', 'Pos.', 'Elig.', 'Height (in.)', 'Hometown / High School')
#reorders column and removes feet and inches columns

```

```{r PSURosterStatsTable}
psuRosterStats <- read_html(
  x = 'https://gopsusports.com/sports/womens-volleyball/stats'
) %>%
  html_elements(css = 'table') %>%
  html_table()
#harvests the player data from the gopsusports website

psuRosterOffensiveHeading <- bind_cols(psuRosterStats[[2]][1, ])
#saves the column headers as a single row

psuRosterOffensiveStats <- bind_cols(psuRosterStats[[2]][2:nrow(psuRosterStats[[2]]) , ])
#saves the rest of the player data as a dataframe


View(psuRosterOffensiveStats)
colnames(psuRosterOffensiveStats) <- c(psuRosterOffensiveHeading[1, ])
#inserts column headers into player data

psuOffensive <- psuRosterOffensiveStats %>%
  select(!contains('Link'))
#removes the 'Bio Link' column

psuOffensive <- psuOffensive[1:17, ]
#removes the 'Total' and 'Opponent' rows

psuOffensive$Player <- str_sub(psuOffensive$Player, end = -25)
#removes the repeated names from the player column

psuOffensive <- psuOffensive %>%
  tidyr::separate_wider_delim(
    cols = Player,
    delim = ',',
    names = c('Last', 'First')
  ) %>%
  #Separates the player names into separate columns for first and last name
  mutate(
    'Player' = paste(First, Last, sep = ' '), 
    .before = First
  ) %>%
  #concatenates the first and last names together
  mutate(
    `Player` = stringr::str_squish(`Player`)) %>%
  select(!'First') %>%
  select(!'Last')
  #removes excess whitespace and removes individual name columns

psuOffensive <- psuOffensive %>%
  mutate_at(vars(!Player), as.numeric)
#turns all columns except Player into numeric types

psuOffensiveRoster <- inner_join(
  x = psuOffensive, 
  y = psuVBRoster, 
  by = join_by('Player' == 'Full Name', '#' == '#')
)
#joins the roster data into the offensive statistics

psuSepPos <- separate_longer_delim(psuOffensiveRoster, 'Pos.', delim = '/')
#creates a second row for players with two positions so each row has only one position

View(psuSepPos)

```

```{r OSURosterStatsTable}
url <- c('https://s3.us-east-2.amazonaws.com/sidearm.nextgen.sites/ohiostatebuckeyes.com/stats/wvb/2023/pdf/cume.pdf')
raw_text <- pdf_text(url)

#the following code was found on the internet to harvest the osu data from a pdf
clean_table <- function(raw) {
  raw <- map(raw, ~ str_split(.x, '\\n') %>% unlist())
  raw <- reduce(raw, c)
  
  table_start <- stringr::str_which(tolower(raw), '#')
  table_end <- stringr::str_which(tolower(raw), '31')
  table_end <- table_end[min(which(table_end > table_start))]
  
  table <- raw[(table_start):(table_end)]
  table <- str_replace_all(table, '\\s{2,}', '|')
  text_con <- textConnection(table)
  data_table <- read.csv(text_con, sep = '|')
  
  data_table
}
osuRawStats <- map_df(raw_text, clean_table)
#the osu data is very dirty, with some columns holding data from more than one column

osuFixOne <- osuRawStats[1, ] %>%
  #the first row has the name and 'SP' in the 'Player' column
  mutate('missing' = paste(Player))
#creates a new column to duplicate the broken cell
osuFixOne$missing <- str_sub(osuFixOne$missing, start = 22)
#removes the name and leaves just the number
osuFixOne$Player <- str_sub(osuFixOne$Player, end = -4)
#removes the number and leaves just the name
osuFixOne <- osuFixOne %>%
  select(!'PTS')
#removes the 'PTS' column
colnames(osuFixOne) <- c('#', 'Player', 'K', 'K.S', 'E', 'TA', 'Pct', 'A', 'A.S', 'SA', 'SE', 'SA.S', 'RE', 'DIG', 'DIG.S', 'BS', 'BA', 'BLK', 'BLK.S', 'BE', 'BHE', 'PTS', 'SP')
#shifts the column headers to reflect the correct data
osuFixOne <- osuFixOne %>%
  select('#', 'Player', 'SP', 'K', 'K.S', 'E', 'TA', 'Pct', 'A', 'A.S', 'SA', 'SE', 'SA.S', 'RE', 'DIG', 'DIG.S', 'BS', 'BA', 'BLK', 'BLK.S', 'BE', 'BHE', 'PTS')
#changes the order of the columns to the original order

osuFixOthers <- osuRawStats[7:12, ] %>%
#rows 7 to 12 include both player number and name in the 'X.' column
  mutate('missing' = paste(X.))
#copies the broken cell into a new column
osuFixOthers$missing <- str_sub(osuFixOthers$missing, start = 4)
#removes the number
osuFixOthers$X. <- str_sub(osuFixOthers$X., end = 3)
#removes the name
osuFixOthers <- osuFixOthers %>%
  select(!'PTS')
#removes the 'PTS' column
colnames(osuFixOthers) <- c('#', 'SP', 'K', 'K.S', 'E', 'TA', 'Pct', 'A', 'A.S', 'SA', 'SE', 'SA.S', 'RE', 'DIG', 'DIG.S', 'BS', 'BA', 'BLK', 'BLK.S', 'BE', 'BHE', 'PTS', 'Player')
#shifts the column names to reflect the correct data


osuGoodRows <- osuRawStats[2:6, ] %>%
  #selects the rows which were formatted correctly
  rename('#' = 'X.')
#changes the name of the player number column

osuStats <- rbind(osuFixOne, osuGoodRows, osuFixOthers)
#vertically joins the osu data back together

osuStats <- osuStats %>%
  tidyr::separate_wider_delim(
    cols = Player,
    delim = ',',
    names = c('Last', 'First')
  ) %>%
  #Separates the player names into separate columns for first and last name
  mutate(
    'Player' = paste(First, Last, sep = ' '), 
    .before = First
  ) %>%
  #concatenates the first and last names together
  mutate(
    `Player` = stringr::str_squish(`Player`)) %>%
  select(!'First') %>%
  select(!'Last') %>%
  #removes the first and last name columns
  mutate_at(vars(!Player), as.numeric)
#saves all columns except for Player as a numeric type

osuJoin <- osuStats %>%
  select('#', 'Player', 'SP', 'K', 'K.S', 'E', 'TA', 'Pct', 'A', 'A.S', 'SA', 'SE', 'SA.S', 'PTS') %>%
  #selects the columns which are in common with the psu data
  mutate('School' = 'OSU')
#adds a column to save the school of each player
colnames(osuJoin) <- c('#', 'Player', 'SP', 'K', 'K/S', 'E', 'TA', 'PCT', 'A', 'A/S', 'SA', 'SE', 'SA/S', 'PTS', 'School')
#changes the names of the columns to match the psu data

psuJoin <- psuOffensive %>%
  select('#', 'Player', 'SP', 'K', 'K/S', 'E', 'TA', 'PCT', 'A', 'A/S', 'SA', 'SE', 'SA/S', 'PTS') %>%
  #selects the columns which are in common with the osu data
  mutate('School' = 'PSU')
#adds a column to save the school of each player
  
psuAndOsu <- rbind(psuJoin, osuJoin)
#vertically joins the psu and osu data together

```

Overall Points Scored Vs. Player Height
```{r PointsVHeight, echo = FALSE}
#Hide the code until the appendix 
ggplot(psuOffensiveRoster) + #Plot based on the PSU offensive roster
  aes(x = `Height (in.)`, y = PTS) + 
  geom_point(shape = "circle", size = 1.8, colour = "#005BAC") +
  labs(
    x = "Height (in.)",
    y = "Points",
    title = "Total Points Vs. Player Height",
    subtitle = "2023 Penn State Womens Volleyball season"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16L),
    axis.title.y = element_text(size = 13L),
    axis.title.x = element_text(size = 13L)
  )

```
The visualization shown above is the comparison of each players height versus their total amount of points for this season. As seen in the graph, the player that has scored the most amount of points this season is 73 in. (6'2 ft) with points ranging in the 500s. The players with the most amount of points range from 73 in (6'2 ft) to 75 in (6'4 ft). We can also see on the graph that the most popular height of the team is 74 in. because they have the most amount of data points on 74in line.


# Conducting Exploratory Data Analysis

```{r averageSetsPlayedPerSchoolYearDataTable}
setsSummaryTable <- psuOffensiveRoster %>%
  # Creates a summary table based on the wrangled and tidied data
  # from the previous code chunk.
  group_by(Elig.) %>%
  # We are grouping by eligibility (grade), meaning that the summary
  # statistics will be given for each grade.
  summarize(
  count = n(),
  minimum = min(SP),
  # Since we are looking specifically at the data for sets played, we select
  # the variable "SP" for all the statistics.
  firstQuintile = quantile(SP, 0.25),
  median = median(SP),
  thirdQuintile = quantile(SP, 0.75),
  maximum = max(SP),
  mean = mean(SP),
  standardDeviation = sd(SP)
    )

setsSummaryTable2 <- setsSummaryTable %>%
  mutate(
    across(where(is.numeric), round, 2)
    )
  # This section utilizes the mutate function to round any variable that
  # is a number to 2 decimal places.

setsSummaryTable2 %>%
  kable(
  col.names = c("Grade","Count", "Minimum", "First Quartile", "Median", "Third Quartile", "Maximum", "Arithmetic Mean", "Arithmetic Standard Deviation"),
  # Changes the column names to those listed above.
  caption = "Summary Statistics of Average Sets Played Per Grade",
  # Changes the title of the summary table.
  booktabs = TRUE,
  align = c("l", rep("c", 6))
  ) %>%
  kableExtra::kable_classic()
  # This section utilizes the kableExtra package to create a clear summary
  # table for the viewer to look at.

View(setsSummaryTable2)
# Mutate function, freshman is first year, etc. 
```
The summary table shown above provides the 5 number summary statistics for number of sets played based on grade (year in school). It also displays the amount of players in each grade, as well as the mean and standard deviation for the number of sets played in each grade. From the summary table, we can see that the freshman had the lowest number of sets played, at 2 sets. The sophomores and seniors had the same maximum number of sets played at 116, though the seniors had a higher mean of total sets played. We can also see that the seniors have the greatest number of starters.

```{r renamingEligibilityRows}
# Reorder function
# reorder(setsSummaryTable2$Grade, c(1, 4, 3, 5, 6, 2))


```

```{r averageSetsPlayedPerSchoolYearDataVisualization}

ggplot(setsSummaryTable2) +
  # Utilizes ggplot to create a data visualization based on the 
  # summary table created previously.
  aes(x = Elig., y = mean, group = count) +
  # Elig. (grade) will be on the horizontal axis, mean will be on the 
  # vertical axis, and the data will overall be grouped by count.
  geom_col(fill = "#3F567E") +
  # geom_col will display a bar chart.
  labs(
    x = "Player Grade (Year)",
    y = "Mean Sets Played",
    title = "Penn State Women's Volleyball Average Number of Sets Played Per Grade (2023 Season)"
  # Sets the axes and title of the data visualization.
  ) +
  theme_minimal()

```
From this data visualization, the viewer can see that the mean number of sets played was lowest for the freshman compared to the other class groups. The mean number of sets played was highest for the seniors, meaning that on average, the seniors played the most sets overall. An important item to note is that the mean number of sets played is higher for the sophomores than it is for the juniors, even though the juniors are older. In this case, the juniors don't play as much as the sophomores, on average, showing that the volleyball team does not choose who plays based on seniority.

```{r averagePointsPerPositionSummaryTable}

statsSummaryTable <- psuSepPos %>%
  # Creates a summary table based on the wrangled and tidied data with object
  # name "psuSepPos" from a previous code chunk.
  group_by(Pos.) %>%
  # We are grouping by position, meaning that the summary
  # statistics will be given for each position.
  summarize(
  count = n(),
  minimum = min(PTS),
  # We are evaluating the number of points scored for each position,
  # which uses the variable PTS.
  firstQuintile = quantile(PTS, 0.25),
  median = median(PTS),
  thirdQuintile = quantile(PTS, 0.75),
  maximum = max(PTS),
  mean = mean(PTS),
  standardDeviation = sd(PTS)
    )

statsSummaryTable2 <- statsSummaryTable %>%
  mutate(
    across(where(is.numeric), round, 2)
    )
  # This section utilizes the mutate function to round any variable that
  # is a number to 2 decimal places.

statsSummaryTable2 %>%
  kable(
  col.names = c("Position","Count", "Minimum", "First Quartile", "Median", "Third Quartile", "Maximum", "Arithmetic Mean", "Arithmetic Standard Deviation"),
  # Changes the column names to those listed above.
  caption = "Summary Statistics of Average Points Earned Per Position",
  # Changes the title of the summary table.
  booktabs = TRUE,
  align = c("l", rep("c", 6))
  ) %>%
  kableExtra::kable_classic()
  # This section utilizes the kableExtra package to create a clear summary
  # table for the viewer to look at.

View(statsSummaryTable2)

```
From the visualization above, one can gather that the number of points depends on where the player is on the court, which depends on their position. For example, while there's only two players who have middle blocker (MB) listed on their position, they have scored the highest average number of points. However, the individual who has the most number of points belongs to the outside hitter (OH), with a maximum of 567.5 points scored. The right side (RS) players also have a relatively high average of points scored, making the three positions with the highest average scores to be the MB, OH, and RS positions. We can see that the Liberos (L) and Defensive Specialist (DS) positions, as well as the setter (S) often don't score. Liberos and Defensive Specialists are often in the rear of the court, which is why they do not score points as much as the right side, outside hitter, and middle blocker, since those are the positions right at the net, where the ball is often spiked. This is depicted in the summary table. 

```{r playerHeightVersusPosition}

ggplot(psuSepPos) +
  # Utilizes ggplot to make a boxplot for the heights compared to the
  # players' positions.
  aes(x = Pos., y = `Height (in.)`) +
  geom_boxplot(fill = "#B0C2E2") +
  labs(
    x = "Player Position",
    y = "Player Height (In.)",
    title = "Penn State Women's Volleyball 2023 Player Height Versus Player Position"
    # Gives the axes labels and title.
  ) +
  theme_minimal()

```
The boxplot above describes the different heights of the different players grouped by their position on Penn State's Women's Volleyball team. The shorter players on the team (depicted by a smaller inch size) are seen to be primarily the Defensive Specialist and the Libero. The tallest players are seen to be Middle Blockers, Outside Hitters, and Right Side. In this way, the tallest players are the ones closer to the net. The taller players have added height that aids them in achieving success in their specific position. Contrarily, the positions of Defensive Specialist and Libero do not necessarily require great height, since they are farther back on the court. The range in height of the Setters (S) varies greatly, showing that many different women, regardless of height, can be successful at that position. 

```{r osuPSUDataComparison}
# Total points
# Average kills per set, service errors per set, who's more consistent 
# Make a graph with both of them on there (red for osu, blue for psu)
# scale_y_binned
# esquisse::esquisser(data = psuAndOsu, viewer = 'browser')

ggplot(psuAndOsu) +
  aes(x = `SA/S`, y = `K/S`, colour = School) +
  geom_point(shape = "circle", size = 1.5) +
  scale_color_manual(
    values = c(OSU = "#E91506",
    PSU = "#1F2987")
  ) +
  labs(
    x = "Kills Per Set Per Player",
    y = "Service Aces Per Set Per Player",
    title = "Service Aces Per Set Per Player Versus Kills Per Set Per Player",
    subtitle = "(2023 Women's Volleyball Statistics for Penn State University and Ohio State University)",
    color = "School"
  ) +
  theme_minimal()

```
The above data visualization depicts an interesting comparison between kills per set per player and aces per set per player, showing if whether those who have more kills also have more aces, and vice versa. A kill is when the player ends the point by hitting the ball into an area that the opposing team cannot get. An ace is when the player serves the ball, but no one on the opposing team is able to receive it. We were aiming to see whether those who had more kills, who are usually more aggressive, since points often come from speed and force of the ball, also used that aggression when serving. Aces would come about from high speed of the ball, though position on the court (where the ball lands/is aimed) is also a contributing factor. As seen in the scatter plot, there is no apparent correlation between aces per set and kills per set per player. Some individuals have a high number of aces, with very few kills, and others have a high number of kills with very few aces, for both teams. Many individuals are clustered around the lower left hand corner of the visualization, showing that they didn't have many aces or kills. When comparing the two different teams (OSU and PSU), one can see that Penn State has two individuals with a higher ace per set rate than the Ohio State players, whereas Ohio State has two individuals with a higher kill per set rate than the majority of Penn State players.

``` {r TotalAttacksPerSet}
Attacks per set versus kills per set (TA/S)

```

## Research Questions

-What is the connection between amount of points a player scores versus their height?
-What are the average stats for each position?
-What is the average number of sets played based on college year?
-How does PSU compare in points scored/number of kills to OSU?

## Read in data and perform necessary wrangling and cleaning
Describe your data sources (where they come from, what was their original purpose, who/what comprise the cases in each).
Notes:
- PSU V. Ohio State Portion: We gathered both the Penn state's and Ohio State's women volleyball team stats. We googled each roster of both teams and merged them together. The PSU stat table did not have the position of each player, so to match each player with their position, we joined a PSU Roster that contained players and their position with the PSU Stat table. We then compared Penn state and Ohio State's stats.

- Only PSU Portion: We used the internet to find the roster and stats of Penn State's Women's Volleyball team. Since the PSU stat roster did not have the position of each player, we merged the stats table with the team roster containing the player positions. We then compared data within the PSU team.

Describe what attributes you'll focus your analysis on (mention if they are part of your data sets or if you created them out of your data sets).

- Visualizations:
- Total points scored vs. height
- Average stats for each position
- Average number of sets played for each college year
- PSU vs. OSU in total points or kills


## Conduct Exploratory Data Analysis




All outputs should be accompanied by narrative text to explain what the reader should be seeing.

Create multiple data visualizations that assist both the team and readers in understanding the data.
Data visualizations should show a variety of your skills and geometries.

Create at least one visual table (not a display of raw data) that assists both the team and readers in understanding the data.

Have narrative text explaining every data visualization and table as well as setting a framework.

