---
title: "PSU Womens Volleyball"
author: "Lily Mayer, Ayla Orona, Kate Miller"
date: "2023-11-29"
output: html_document
---

=======

All code should be written according to a Style Guide of your choice. List this Style Guide as a code comment in your first code chunk.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# List Style Guide in this code chunk
library(rvest)
library(dplyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(stringr)
```

```{r espnTable}
espnPSUTables <- read_html(
  x = "link"
  # I haven't found any ESPN related items, so we may have to rethink where we get data from.
) %>%
  html_elements(css = "table") %>%
  html_table()
View(espnPSUTables)
```

```{r PSURosterTable}
psuVBRoster <- read_html(
  x = "https://gopsusports.com/sports/womens-volleyball/roster"
) %>%
  html_elements(css = "table") %>%
  html_table()
psuVBRoster <- bind_cols(psuVBRoster[[3]])
View(psuVBRoster)
```

<<<<<<< HEAD
```{r PSURosterStatsTable}
psuRosterStats <- read_html(
  x = 'https://gopsusports.com/sports/womens-volleyball/stats'
) %>%
  html_elements(css = 'table') %>%
  html_table()
psuRosterOffensive <- bind_cols(psuRosterStats[[2]])

View(psuRosterOffensive)
```
=======
```{r OSURosterTable}
osuVBRoster <- read_html(
  x = "https://ohiostatebuckeyes.com/sports/womens-volleyball/roster"
) %>%
  html_elements(css = "table") %>%
  html_table()
View(osuVBRoster)
```

## Research Questions
Still deciding...
based on summary table of PSU team and visualizations

## Read in data and perform necessary wrangling and cleaning
Describe your data sources (where they come from, what was their original purpose, who/what comprise the cases in each).
Notes:
- PSU V. Ohio State Portion: We gathered both the Penn state's and Ohio State's women volleyball team stats. We googled each roster of both teams and merged them together. The PSU stat table did not have the position of each player, so to match each player with their position, we joined a PSU Roster that contained players and their position with the PSU Stat table. We then compared Penn state and Ohio State's stats.

- Only PSU Portion: We googled the roster and stats of the Penn state's women volleyball team. Because the PSU stat roster did not have the position of each player, we merged the stats table with the team roster containing players position. We then compared data within the PSU team.

Describe what attributes you'll focus your analysis on (mention if they are part of your data sets or if you created them out of your data sets).

-service ace per set vs service error per set. To get rate of errors: SE column divided by SP column.
- Summary Table: average height, kills, digs, etc. possibly for each position or as a whole
- Visualizations: stats of each year for each player(bar chart, histogram), PSU v. OSU, etc.

## Conduct Exploratory Data Analysis
All outputs should be accompanied by narrative text to explain what the reader should be seeing.

Create multiple data visualizations that assist both the team and readers in understanding the data.
Data visualizations should show a variety of your skills and geometries.

Create at least one visual table (not a display of raw data) that assists both the team and readers in understanding the data.

Have narrative text explaining every data visualization and table as well as setting a framework.
