---
title: "PSU Womens Volleyball"
author: "Lily Mayer, Ayla Orona, Kate Miller"
date: "2023-11-29"
output: html_document
---

All code should be written according to a Style Guide of your choice. List this Style Guide as a code comment in your first code chunk.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# List Style Guide in this code chunk
```

```{r espnTable}
espnPSUTables <- read_html(
  x = "link"
  # I haven't found any ESPN related items, so we may have to rethink where we get data from.
) %>%
  html_elements(css = "table") %>%
  html_table()
View(espnPSUTables)
```

```{r PSURosterTable}
psuVBRoster <- read_html(
  x = "https://gopsusports.com/sports/womens-volleyball/roster"
) %>%
  html_elements(css = "table") %>%
  html_table()
View(psuVBRoster)
```

## Research Questions
State your research questions and explain them.

## Read in data and perform necessary wrangling and cleaning
Describe your data sources (where they come from, what was their original purpose, who/what comprise the cases in each).

Describe what attributes you'll focus your analysis on (mention if they are part of your data sets or if you created them out of your data sets).

## Conduct Exploratory Data Analysis
All outputs should be accompanied by narrative text to explain what the reader should be seeing.

Create multiple data visualizations that assist both the team and readers in understanding the data.
Data visualizations should show a variety of your skills and geometries.

Create at least one visual table (not a display of raw data) that assists both the team and readers in understanding the data.

Have narrative text explaining every data visualization and table as well as setting a framework.
